---
title: "PA1_template"
author: "Aaron Taylor"
date: "September 14, 2017"
output: html_document
---

******
#Reproducible Research: Course Project 1
******
##Retrieve File from Internet + Preprocessing
###Downloads from internet and unzips
```{r loaddata}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
temp <- tempfile() #No need to store uncessary .zip files
download.file(fileUrl, destfile = temp, mode = "wb") #add a ", method = "curl"" if using a mac
dateDowloaded <- date()
dateDowloaded
unzip(temp, "activity.csv") #unzips the file, turns it into a readable .txt file
ActivityData <- read.csv("activity.csv", header = TRUE)
unlink(temp)
```

###Converts Factor-class dates into Date-class dates
```{r}
library(lubridate)
ActivityDataDate <- ActivityData
ActivityDataDate$date<-ymd(ActivityData$date)
```

###Cleans the data of NA's
```{r}
ActivityDataDateClean <- na.omit(ActivityDataDate)
```

##What is mean total number of steps taken per day?
###Totals per day calculation
```{r}
totalStepsPerDay <- c()
numDates <- length(unique(ActivityDataDateClean$date))
for(i in 1:numDates) {
    dates <- unique(ActivityDataDateClean$date)[i]
    totalStepsPerDay[i] <-
        sum(ActivityDataDateClean$steps[ActivityDataDateClean$date == dates])
}
```

###Create the "total number of steps" histogram
```{r}
hist(totalStepsPerDay, xlab = "Total Steps Per Day", main = "Total Steps Per Day")
```

###Mean and median of the total
```{r}
mean(totalStepsPerDay)
median(totalStepsPerDay)
```

******

##What is the average daily activity pattern?
###Mean per Interval calculation
```{r}
meanStepsPerInterval <- c()
numIntervals <- length(unique(ActivityDataDateClean$interval))
for(i in 1:numIntervals) {
    intervals <- unique(ActivityDataDateClean$interval)[i]
    meanStepsPerInterval[i] <-
        mean(ActivityDataDateClean$steps[ActivityDataDateClean$interval == intervals])
}
```

###Create the "mean number of steps" plot
```{r}
plot(x = unique(ActivityDataDateClean$interval), y = meanStepsPerInterval,
     type = "l", xlab = "Interval (Minutes)", ylab = "Mean Steps",
     main = "Daily Activity Pattern")
```

###Interval with the maximum number of steps
```{r}
unique(ActivityDataDateClean$interval)[which(meanStepsPerInterval == max(meanStepsPerInterval))]
```
*Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?*  
Interval 835, which I think is 8:35 AM, or 08:35.  

******

##Imputing missing values
###Find # rows with NAs
```{r}
nrow(ActivityDataDate[is.na(ActivityDataDate),])
```

###Fill the NA's with the mean-per-interval; aggregate() makes this easy
```{r}
ActivityDataNAFilled <- as.data.frame(ActivityDataDate)
ActivityDataAggregate <- aggregate(steps ~ interval, ActivityDataDate, mean)
ActivityDataNAFilled$steps[is.na(ActivityDataNAFilled$steps)] <- ActivityDataAggregate$steps
```

###Imputed total number of steps calculation
```{r}
totalStepsPerDayImputed <- c()
numDates1 <- length(unique(ActivityDataNAFilled$date))
for(i in 1:numDates1) {
    dates1 <- unique(ActivityDataNAFilled$date)[i]
    totalStepsPerDayImputed[i] <- sum(ActivityDataNAFilled$steps[ActivityDataNAFilled$date == dates1])
}
```

###Imputed total number of steps histogram plotter
```{r}
hist(totalStepsPerDayImputed, xlab = "Total Steps Per Day", main = "(Imputed) Total Steps Per Day")
```

###Mean and median of the imputed total
```{r}
mean(totalStepsPerDayImputed)
median(totalStepsPerDayImputed)
```
*Do these values differ from the estimates from the first part of the assignment?*  
The mean is the same; the median has changed only slightly, by ~0.01%.  
*What is the impact of imputing missing data on the estimates of the total daily number of steps?*  
The central spire, 10000-15000, has gotten higher, as one might expect from putting extra weight on the mean.

******

##Are there differences in activity patterns between weekdays and weekends?
###sets up two new columns, for day-of-week and weekend/day type
```{r}
ActivityDataNAFilledDay <- as.data.frame(ActivityDataNAFilled)
ActivityDataNAFilledDay$day <- weekdays(ActivityDataNAFilled$date)
ActivityDataNAFilledDay$weekday <- weekdays(ActivityDataNAFilled$date)
```

###Sorts the rows by weekend or weekday
```{r}
numDates2 <- length(ActivityDataNAFilledDay$date)
for(i in 1:numDates2) {
    if (ActivityDataNAFilledDay$day[i] == "Saturday" || ActivityDataNAFilledDay$day[i] == "Sunday") {
        ActivityDataNAFilledDay$weekday[i] <- "Weekend"
    }
    else {
        ActivityDataNAFilledDay$weekday[i] <- "Weekday"
    }
}
```

###Convert weekday from character to factor, needed for the aggregate
```{r}
ActivityDataNAFilledDay$weekday <- as.factor(ActivityDataNAFilledDay$weekday)
```

###Aggregate once again proves useful; use steps as interval to get average number of steps in an interval across all days
```{r}
ADNAFDInterval <- aggregate(steps ~ interval+weekday, ActivityDataNAFilledDay, mean)
```

###Make the panel plot for weekdays/ends
```{r}
library(ggplot2)
qplot(interval, steps, data=ADNAFDInterval, geom=c("line"), xlab="Interval", ylab="Number of steps", main="") + facet_wrap(~ weekday, ncol=1)
```

******

##Removes now-unnecessary data from memory
```{r}
rm(ActivityData, ActivityDataAggregate, ActivityDataDate, ActivityDataDateClean, ActivityDataNAFilled, ActivityDataNAFilledDay, ADNAFDInterval)
```